---
title: "CS 422 HW2"
output: 
  html_notebook:
    toc: yes
    toc_float: yes
author: Matthew Venter
---

### Loading librarys
```{r}
library(dplyr)
library(plyr)
library(rpart)
library(caret)
us.train$Scaled <- as.factor(us.train$Scaled)
library(ROCR)
library(randomForest)

```

### Part 2.1-a 
```{r}
hotel.df <- read.csv("hotel_bookings.csv", sep=",", header=T)
hotel.df$is_canceled <- as.factor(hotel.df$is_canceled)
hotel.df$same_room <- as.character(hotel.df$reserved_room_type)== as.character(hotel.df$assigned_room_type)#creating a varible that tracks if a person asked for a type of room and got the same type.
cat("Resort hotel frequency:", sum(hotel.df$hotel=="Resort Hotel"),"\nCity hotel frequency:", sum(hotel.df$hotel=="City Hotel"))
```
### Part 2.1-b
```{r}
cat("Number of guests who canceled reservation:",sum(hotel.df$is_canceled==1) ,"\nNumber of guests did not cancel the reservation:",sum(hotel.df$is_canceled==0))
```
### Part 2.1-c
```{r}
cust.type.df <- count(hotel.df,"customer_type")
cat("Customer type with the most reservations is Transient with", cust.type.df[which.max(cust.type.df$freq),2],"reservations")

```
### Part 2.1-d
```{r}
var0 <- hotel.df[which.max(hotel.df$required_car_parking_spaces),"required_car_parking_spaces"]
cat(sum(hotel.df$required_car_parking_spaces == var0),"customers required the most number of parking spaces(" ,var0,")")
```
### Part 2.1-e
```{r}
var1 <- hotel.df[which.min(hotel.df$required_car_parking_spaces),"required_car_parking_spaces"]
cat(sum(hotel.df$required_car_parking_spaces == var1),"customers required the least number of parking spaces(" ,var1,")")

```
### Part 2.1-f
```{r}
perc_good <- sum(as.character(hotel.df$reserved_room_type)==as.character(hotel.df$assigned_room_type))/119390*100
cat(perc_good,"% of the people who expressed a room preference during reservation got the room during check-in.")

```
### Part 2.1-g
```{r}
hist.hotel.info.h1 <- filter(hotel.df, hotel == "City Hotel") 
hist.hotel.info.h2 <- filter(hotel.df, hotel == "Resort Hotel"& country!= "NULL")
hist.hotel.table.h1 <- table(hist.hotel.info.h1$country)
hist.hotel.table.h2 <- table(hist.hotel.info.h2$country)
barplot(tail(sort(hist.hotel.table.h1),n=10), main = "Top 10 Origins for Resort Hotels",ylab = "Amount of bookings per country",col = rainbow(10))
barplot(tail(sort(hist.hotel.table.h2),n=10),main = "Top 10 Origins for City Hotels",ylab = "Amount of bookings per country",col = rainbow(10))
```
### Part 2.1-h
 i.) The country with the most origins of either type of hotel was Portugal <br />
ii.) We can conclude that the hotel information was most likely taken from hotels in portugal or nearby

### Part 2.2-a
```{r}
#a.)
set.seed(1122)
index <- sample(1:nrow(hotel.df), size=0.1*nrow(hotel.df))
hotel.test <- hotel.df[index, ]
hotel.train <- hotel.df[-index, ]
hotel.model <- rpart(is_canceled ~ deposit_type + lead_time +country + total_of_special_requests + required_car_parking_spaces + agent +same_room , method="class", data=hotel.train)
```

### Part 2.2.a.i
```{r}
rpart.plot(hotel.model, type = 4, extra = 4,branch.lty = 7, box.palette = "RdYlGn",compress =TRUE, ycompress=TRUE,  main="Hotel Cancelation Model")
```

### Part 2.2.a.ii
```{r}
print("The varibibles I chose were deposit_type, lead_time, country, total_of_special_requests, required_car_parking_spaces, + agent, and same_room")
```

### Part 2.2.a.iii
```{r}
hotel.fit <- predict(hotel.model, hotel.test, type = "class")
confusionMatrix(hotel.fit,hotel.test$is_canceled)
```

### Part 2.2.a.iv
```{r}
pruned.rocr <- predict(hotel.model, newdata=hotel.test, type="prob")[,2]
f.pred <- prediction(pruned.rocr, hotel.test$is_canceled)
plot(performance(f.pred, "tpr", "fpr"), colorize=T, lwd=3)
abline(0,1)
```

#### Part 2.2.a.v
```{r}
auc.pruned <- performance(f.pred, measure = "auc")
cat(paste("The area under curve (AUC) for the full tree is ", 
          round(auc.pruned@y.values[[1]], 3)))
```

#### Part 2.2.b.i
```{r}
printcp(hotel.model)
plotcp(hotel.model)
```
#### Part 2.2.b.ii
To me it seemed that the tree did not need pruning as there isnt a point in which it becomes more inacurate as it keeps learning (model may be underfitted)

#### Part 2.3.a.i
```{r}
hotel.train$country = as.numeric(hotel.train$country)
hotel.train$agent = as.numeric(hotel.train$agent)
hotel.test$country = as.numeric(hotel.test$country)
hotel.test$agent = as.numeric(hotel.test$agent)
model.rf.mtry4 <- randomForest(is_canceled ~ deposit_type + lead_time  + total_of_special_requests + required_car_parking_spaces + country + agent + same_room, data = hotel.train, mtry=4, ntree=750, importance=TRUE, do.trace=250)
rf.predict <- predict(model.rf.mtry4, hotel.test, type="class")
confusionMatrix(rf.predict, hotel.test$is_canceled)
model.rf.mtry6 <- randomForest(is_canceled ~ deposit_type + lead_time  + total_of_special_requests + required_car_parking_spaces + country + agent, data = hotel.train, mtry=6, ntree=750, importance=TRUE, do.trace=250)
rf.predict <- predict(model.rf.mtry6, hotel.test, type="class")
confusionMatrix(rf.predict, hotel.test$is_canceled)
model.rf.mtry10 <- randomForest(is_canceled ~ deposit_type + lead_time  + total_of_special_requests + required_car_parking_spaces + country + agent, data = hotel.train, mtry=10, ntree=750, importance=TRUE, do.trace=250)
rf.predict <- predict(model.rf.mtry10, hotel.test, type="class")
confusionMatrix(rf.predict, hotel.test$is_canceled)


```
#### Part 2.3.a.i
```{r}
# 500,4 is the best!!!!!!!!!!!!!!!!!!!!!!!!
model.rf <- randomForest(is_canceled ~ deposit_type + lead_time  + total_of_special_requests + required_car_parking_spaces + country + agent, ntree = 500,    mtry = 4, data = hotel.train)
rf.predict <- predict(model.rf, hotel.test, type="class")
confusionMatrix(rf.predict, hotel.test$is_canceled)
print("The best prediction was with ntree = 500 and mtry = 4 it had an Accuracy of 85.33%, Sensitivity of 0.9048, Specificity of 0.7703, and Balanced Accuracy of 0.8376")
```
#### Part 2.3.a.ii
```{r}
# 500,4 is the best also for oob with a value of 14.60%
model.rf <- randomForest(is_canceled ~ deposit_type + lead_time  + total_of_special_requests + required_car_parking_spaces + country + agent + same_room, ntree = 500, mtry = 4, data = hotel.train)
rf.predict <- predict(model.rf, hotel.test, type="class")
confusionMatrix(rf.predict, hotel.test$is_canceled)
print("The best prediction was with ntree = 500 and mtry = 4 it had an Accuracy of 85.61%, Sensitivity of 0.9060, Specificity of 0.7743, and Balanced Accuracy of 0.8402")
```
#### Part 2.3.a.iii
They are the same model I think this makes sense becasue with too many trees it starts over fitting so the accuracy will go down and the OOB wil go up





