---
title: "CS 422 HW3"
output: 
  html_notebook:
    toc: yes
    toc_float: yes
author: Matthew Venter
---

### Loading librarys
```{r}
library(dplyr)
library(plyr)
library(arules)
library(arulesViz)
library(cluster)
library(factoextra)
library(fpc)
```

### Part 2.1-a Creating data frames needed
```{r}
#creating regular dataframes
tr1k.df<- read.transactions(file ="Tr_1k_can.csv", sep=",", header=F)
tr5k.df<- read.transactions(file ="tr-5k-can.csv", sep=",", header=F)
tr20k.df<- read.transactions(file ="tr-20k-can.csv", sep=",", header=F)
tr75k.df<- read.transactions(file ="tr-75k-can.csv", sep=",", header=F)
```
### Part 2.1-b Making rules
```{r echo = T, results = 'hide'}
tr1k.freq <- apriori(tr1k.df, parameter=list(support=0.0455, conf = 0.1, target="frequent itemsets"))
tr1k.rules <- apriori(tr1k.df, parameter=list(support=0.0455, conf = 0.1))

tr5k.freq <- apriori(tr5k.df, parameter=list(support=0.0455,conf = 0.1, target="frequent itemsets"))
tr5k.rules <- apriori(tr5k.df, parameter=list(support=0.0455, conf = 0.1))

tr20k.freq <- apriori(tr20k.df, parameter=list(support=0.0445,conf = 0.1, target="frequent itemsets"))
tr20k.rules <- apriori(tr20k.df, parameter=list(support=0.0445, conf = 0.1))

tr75k.freq <- apriori(tr75k.df, parameter=list(support=0.0438,conf = 0.1))
tr75k.rules <- apriori(tr75k.df, parameter=list(support=0.0438,conf = 0.1, target="frequent itemsets"))
```

### Part 2.1-b Evaluating rules
```{r}
#tr1k inspection
inspect(sort(tr1k.freq, decreasing = T, by="count"))
inspect(sort(tr1k.rules, decreasing = T, by="count"))
#tr5k inspection
inspect(sort(tr5k.freq, decreasing = T, by="count"))
inspect(sort(tr5k.rules, decreasing = T, by="count"))
#tr20k inspection
inspect(sort(tr20k.freq, decreasing = T, by="count"))
inspect(sort(tr20k.rules, decreasing = T, by="count"))
#tr75k inspection
inspect(sort(tr75k.freq, decreasing = T, by="count"))
inspect(sort(tr75k.rules, decreasing = T, by="count"))
#del(tr1k.freq)
#del(tr1k.rules)
```
### Part 2.1-c Evaluating rules
As there are more and more observations the support for rules decrease slightly, the confidence raises with more obsevations.(count also increases but this is obvious)

### Part 2.1-d Evaluating questions given
i.) The most frequently purchaed items in 75k transations was {coffee eclair} with a count of 8193
ii.) For the support given the least purchased items are {Gongolais Cookie,Truffle Cake} with a count of 3294

### Part 2.2-a Data cleanup
```{r}
#creating data frame for mammels
mammals.df <- read.csv("mammels.csv", sep=",", row.name = 1, header=T)
```
i.) I didn't want to remove any of the atributes as they all seemed important and not every top tooth type had the same amount of bottom teeth types.
ii.) Data already seemed fine to work with as there are no crazy outliers and the range of values are fairly close together 
iii.) I moved it to a csv file above!

### Part 2.2-b Clustering 
```{r}
#i.)determining optimal number of clusters
fviz_nbclust(mammals.df[1:8], kmeans, method="wss")
fviz_nbclust(mammals.df[1:8], kmeans, method="silhouette")

#ii.)using optimal number of clusters
mammals.kmeans <- kmeans(mammals.df[1:8], centers=8, nstart=25)
fviz_cluster(mammals.kmeans, data=mammals.df[1:8])

#iii.) how many observations?
cat("Number of Observationssum in cluster 1: ",sum(mammals.kmeans$cluster==1))
cat("Number of Observationssum in cluster 2: ",sum(mammals.kmeans$cluster==2))
cat("Number of Observationssum in cluster 3: ",sum(mammals.kmeans$cluster==3))
cat("Number of Observationssum in cluster 4: ",sum(mammals.kmeans$cluster==4))
cat("Number of Observationssum in cluster 5: ",sum(mammals.kmeans$cluster==5))
cat("Number of Observationssum in cluster 6: ",sum(mammals.kmeans$cluster==6))
cat("Number of Observationssum in cluster 7: ",sum(mammals.kmeans$cluster==7))
cat("Number of Observationssum in cluster 8: ",sum(mammals.kmeans$cluster==8))
#iv.) 
cat("Total SSE:",mammals.kmeans$tot.withinss)
#v.)
cat("SSE for each cluster:",mammals.kmeans$withinss)
```
vi.)Analyisis Looking at the clustering and how the mammels fit in a species I can determine that the clustering does make sense and is correct

### Part 2.3-a 
```{r}
set.seed(1122)
samp.mam <- mammals.df[sample(nrow(mammals.df), 35), ]
samp.mam<-scale(samp.mam)
mammal.hc.single <- hclust(dist(samp.mam),method="single")
mammal.hc.complete <- hclust(dist(samp.mam),method="complete")
mammal.hc.avg <- hclust(dist(samp.mam),method="average")
fviz_dend(mammal.hc.single,main = "Cluster Dendogram Single")
fviz_dend(mammal.hc.complete, main = "Cluster Dendogram Complete")
fviz_dend(mammal.hc.avg, main = "Cluster Dendogram AVG")
```
### Part2.3-b
```{r}
print("Single method clustering has the following two singletons: {Elk, Raindeer}, {Sea Lion, Fur Seal}, {Lump Nose Bat,Brown Bat} for a total of 3")
print("Complete method clustering has the following two singletons: {Elk, Raindeer}, {Sea Lion, Fur Seal}, {Lump Nose Bat,Brown Bat}, {House Bat, Red Bat}, {Star Nose Mole, Raccoon} for a total of 5")
print("Avg method clustering has the following two singletons: {Elk, Raindeer}, {Sea Lion, Fur Seal}, {Lump Nose Bat,Brown Bat}, {House Bat, Red Bat}, {Star Nose Mole, Raccoon} for a total of 5")
```

### Part2.3-b
```{r}
print("I would consider the single method to be the most pure method")
```

### Part2.3-d
```{r}
fviz_dend(mammal.hc.single,main = "Cluster Dendogram Single",h=2)
print("If you were to draw a line at h=2 it would result in 7 differnt clusters (each color represents a cluster")
```
###Part 2.3-e
```{r}
fviz_dend(mammal.hc.single,main = "Cluster Dendogram Single",k=7)
fviz_dend(mammal.hc.complete, main = "Cluster Dendogram Complete",k =7)
fviz_dend(mammal.hc.avg, main = "Cluster Dendogram AVG",k =7)
```
###Part 2.3-f
```{r}
mammal.hc.single <- eclust(samp.mam, "hclust", hc_method = "single")
mammal.hc.complete <- eclust(samp.mam, "hclust", hc_method ="complete")
mammal.hc.avg <- eclust(samp.mam, "hclust", hc_method ="average")
mammal.hc.single.stats <-cluster.stats(dist(samp.mam),mammal.hc.single$cluster)
mammal.hc.complete.stats <-cluster.stats(dist(samp.mam),mammal.hc.complete$cluster)
mammal.hc.avg.stats <-cluster.stats(dist(samp.mam),mammal.hc.avg$cluster)
cat("avg silwidth of single hclust :",mammal.hc.single.stats$avg.silwidth)
cat("dunn index of single hclust :",mammal.hc.single.stats$dunn)
cat("avg silwidth of complete hclust :",mammal.hc.complete.stats$avg.silwidth)
cat("dunn index of complete hclust :",mammal.hc.complete.stats$dunn)
cat("avg silwidth of avg hclust :",mammal.hc.avg.stats$avg.silwidth)
cat("dunn index of avg hclust :",mammal.hc.avg.stats$dunn)
```
### Part2.3-g
<br/>
The best linkage strategy according to sillhouette is "single"
<br/>The best linkage strategy according to dunn index is "single"









